### 2015-08-24 TeXFiles/Manuscript_Albert_Ulzega.tex
draft for internal review

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent
% ( open parenthesis    ) close paren.
% - hyphen              = equals sign
% | vertical bar        ~ tilde
% @ at sign             _ underscore
% { open curly brace    } close curly
% [ open square         ] close square bracket
% + plus sign           ; semi-colon
% * asterisk            : colon
% < open angle bracket  > close angle
% , comma               . full stop
% ? question mark       / forward slash
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ
% abcdefghijklmnopqrstuvwxyz
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt,a4paper,final]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP journals}}
%Uncomment next line if AMS fonts required
\usepackage{iopams}
\usepackage{graphicx}
\usepackage[breaklinks=true,colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=blue]{hyperref}
\usepackage{url}
\usepackage{caption}
%\usepackage{amsmath,amssymb,amsthm,amscd}
%\usepackage{natbib}

\usepackage[all]{xy}
\newcommand{\eq}{\mbox{eq}}
\newcommand{\dt}{\mathit{dt}}
\newcommand{\bbM}{\mathbb M}
\newcommand{\bt}{\pmb\theta}
\newcommand{\Bt}{\pmb\Theta}
\newcommand{\vc}[1]{{\bf #1 }}

\begin{document}

\title[Parameter inference with stochastic models]{Bayesian Parameter Inference for 1D Nonlinear Stochastic Differential Equation Models}

\author[cor1]{Carlo Albert}%$^{1,2}$
\address{Eawag, Swiss Federal Institute of Aquatic Science and Technology, 8600 D\"ubendorf, Switzerland}
%\address{$^2$Address Two, Neverland}
\ead{carlo.albert@eawag.ch}

\author{Simone Ulzega}
\address{Eawag, Swiss Federal Institute of Aquatic Science and Technology, 8600 D\"ubendorf, Switzerland}
\ead{simone.ulzega@eawag.ch}

%\author[cor1]{Author Three}
%\address{Address Four, Neverland}
%\eads{\mailto{author.three@mail.com}, \mailto{author.three@gmail.com}}

\begin{abstract}
Bayesian statistics has become an indispensable tool in many applied sciences, in particular, for the purpose of uncertainty analysis.
Inferring parametric uncertainty, for stochastic differential equation (SDE) models, however, is a computationally hard problem due to the high dimensional integrals that have to be calculated.
%Here we present a Monte Carlo-based method for Bayesian parameter inference with generic one dimensional SDE models to tackle the generic problems of calibrating the model to given data time series while simultaneously quantifying the ensuing parametric uncertainty.
Here, we consider the generic problem of calibrating a one dimensional SDE model to time series and quantifying the ensuing parametric uncertainty.
We re-interpret the Bayesian posterior distribution, for model parameters, as the partition function of a statistical mechanical system and employ a Hamiltonian Monte Carlo algorithm to sample from it.
Depending on the number of discretization points and the number of measurement points the dynamics of this system happens on very different time scales.
Thus, we employ a multiple time scale integration together with a suitable re-parametrization to derive an efficient inference algorithm.
While the algorithm is presented by means of a simple SDE model from hydrology, it is readily applicable to a wide range of inference problems. Furthermore the algorithm is highly parallelizable.

\end{abstract}

%Uncomment for PACS numbers title message
\pacs{00.00, 20.00, 42.10}
% Keywords required only for MST, PB, PMB, PM, JOA, JOB?
\vspace{2pc}
\noindent{\it Keywords}: Bayesian parameter inference, Hamiltonian Monte Carlo, time series analysis, stochastic differential equations, path-integrals

% Uncomment for Submitted to journal title message
\submitto{\NJP}
% Comment out if separate title page not required
%\maketitle

\section{Introduction}

Phenomenological models are employed in many applied areas of research to predict the behaviour of complex systems of all sorts.
Parameters of these models often don't have a direct physical meaning and need to be inferred from measurements.
In order to make reliable predictions with such models it is important to describe the dominant errors in the model and to quantify the parametric uncertainty resulting from the inference process.

{\em Bayesian statistics} describes knowledge about parameters through probability distributions and model-based learning through a consistent update rule. It is thus well suited to quantify parametric uncertainty resulting from a model based inference process.
Bayesian statistics is commonly used in many applied sciences and is growing in importance in the physics community as well \cite{vonToussaint_2011}.

A faithful description of the dominant errors in a model naturally leads to {\em stochastic differential equations} (SDEs).
The kind of problems we consider here are the calibration of ordinary 1D SDE models to noisy time series and the quantification of the resulting parametric uncertainty.
Whilst the techniques we use are generic they will be presented by means of a simple yet non trivial SDE model from hydrology.

Bayesian parameter inference for SDE models is computationally very expensive, as the posterior probability density for the parameters is a {\em path-integral}.
Problems of this kind are commonly solved by means of {\em Monte Carlo} methods that are based on simulating model realizations and comparing them to the data. Algorithms of this kind are {\em particle filters} \cite{chopin_2013_SMC2}. For the case of linear SDEs that are coupled to nonlinear deterministic ordinary differential equations (ODEs), more efficient algorithms of this kind can be derived \cite{tomassini_2009_smoothing, reichert_2009_timedepParameters}.
The problem with these simulation-based methods is their inefficiency in the presence of many data points.
One solution is to map the output space to a smaller dimensional space of {\em summary statistics}, and acccept/reject proposed model parameters depending on how compatible associated model runs are with the data in terms of these summary statistics.
Such {\em Approximate Bayes Computations} \cite{marin_2012_ABC, Albert_2014_ABC} are relatively easy to apply as they only require us to run the simulator. However, it is largely an unsolved problem how to choose the summary statistics so as to achieve a sufficient approximation of the posterior parameter distribution.

Exact inference algorithms of high efficiency can be derived from a reinterpretation of the posterior distribution as the partition function of a statistical mechanical system and by simulating the dynamics of the latter.
Upon discretizing time of the original problem we are led to the statistical mechanics of a polymer with harmonic bonds in an exterior potential. The measurements are interpreted as an additional exterior potential acting on the measurement beads of the polymer only.
Furthermore, model parameters are interpreted as additional degrees of freedom coupling to all the beads of the polymer.
To simulate the dynamics of this system we apply the {\em Hamiltonian Monte Carlo} (HMC) algorithm \cite{duane_1987}, which combines Molecular Dynamics (MD) \cite{alder_1959_MD, rahman_1964_MD} with the Metropolis algorithm \cite{metropolis_1953}.
Compared to filtering methods, we achieve much higher acceptance rates since data points are already used for the suggestion of new parameters, and thus model realizations that are incompatible with the data are never even considered.
The drawback of these methods is that the model equations need to be known and derivatives have to be calculated. The latter problem, however, is largely remedied by the use of automated differentiation algorithms.

Tuning of HMC algorithms is a non-trivial matter.
Two sorts of tuning parameters need to be adjusted: (i) those of the {\em kinetic energy} of the statistical mechanical system and (ii) those of the {\em numerical integration} scheme of Hamilton's equation in the molecular dynamics part of the HMC algorithm.
It has been shown that efficiency is gained when the kinetic term is made dependent on the configuration of the statistical mechanical system \cite{girolami_2011_HMC}.
Here we explore a different, computationally simper route.
Depending on the number of discretization points needed to approximate the original SDE system and the number of measurement points, the dynamics of the statistical mechanical system happens on very different time scales.
Thus, we employ a multiple time scale integration technique for the simulation of the statistical mechanical system \cite{tuckerman_1993}.
At least for 1D SDEs we always find a parametrization, which, to some extent, decouples harmonic modes in between measurement points from both the measurement points and the model parameters, and allows for an efficient analytical solution of their dynamics.

The resulting algorithm is very efficient and, through the use of automated differentiation, easily applicable to a wide range of inference problems with SDEs.
Furthermore, it is easily parallelizable.

\section{An Exemplary Inference Problem}

The methods we present are generic and applicable to a wide range of SDE models.
However, with later applications in mind and to make things concrete, we present them by means of a simple yet non-trivial model from hydrology.
Consider, for simplicity, a hydrological catchment whose dynamics at the observation time scale is well described by a linear reservoir and whose other processes happen at much shorter time scales so that they can be described by white noise.
Furthermore, assume this noise to scale linearly with the system state, $S(t)$, which is the water content in the reservoir.
The model equation is thus given by the SDE
\begin{equation}\label{sde}
\dot{S}(t) = r(t) - \frac{1}{K}\left(1+\frac{\gamma}{2}\right) S(t)
+
\sqrt{\frac{\gamma}{K}} S(t){\eta}(t)\,,
\end{equation}
where $r(t)$ denotes the time varying rain input and $\eta(t)$ denotes white noise, i.e.,
\begin{equation}\label{whitenoise}
\langle\eta(t)\eta(t')\rangle = \delta(t-t')\,.
\end{equation}
Eq.~(\ref{sde}) is to be understood in the Stratonovich sense~\cite{stratonovich_1968}.
Our parametrization is such that, for constant rain input $r(t)=r_0$ and in the long-time limit, the mean of $S(t)$ converges to the equilibrium solution of the unperturbed ($\gamma=0$) system, $S_{\eq}=Kr_0$ (see Eq.~(\ref{equilibrium_mean}) below).
Scale-invariance of eq. (\ref{sde}), for large $S$, leads to power law tails in the system state probability distribution (see Eq.~(\ref{inverse_gamma})), which is in line with the observation that errors in hydrological models are often fat-tailed~\cite{thyer_2009_fattails}.

Conceptual models of this kind are commonly used in hydrology, for the purpose of predicting rainfall-runoff behaviour, for natural and urban catchments (see, e.g., \cite{breinholt_2011_SDE}).
Furthermore, by means of the transformation $S(t)=1/n(t)$, eq. (\ref{sde}) turns into a model that has been suggested as a phenomenological description of the dynamics of the neutron density in nuclear reactors \cite{dutre_1977_SDE} and extensively studied, from a theoretical point of view \cite{fujisaka_1986_intermittency}.

Here, we consider the input $r(t)$ to be a smooth and nowhere vanishing function. Realistic multi-fractal rain inputs \cite{tessier_1996} will be studied elsewhere.
Indeed, the focus of this work is not on hydrological modeling, but on generic methods for parameter inference with nonlinear SDE models that are calibrated to observed time-series.
Here we assume the observed time-series, $y_s$, to be the outflow of the reservoir, $S(t)/K$, observed at times $0=t_1<t_2<\dots < t_{n+1}=T$, with multiplicative independent log-normal errors, i.e.,
\begin{equation}\label{data}
  \ln \left( y_s \right)
  =
  \ln \left( \frac{S(t_s)}{K} \right)
  +
  \sigma\epsilon_s\,,\quad s=1,\dots,n+1\,,
\end{equation}
where the $\epsilon_i$ are uncorrelated standard normal errors.

For simplicity, we assume $\sigma$ as well as the input $r(t)$ to be known so that we are left with the task of inferring parameter combinations $(K,\gamma)$ that are compatible with the data given by Eq.~(\ref{data}).
Here, "compatible" is meant in the Bayesian sense, in which knowledge about parameters, $\bt$, is expressed in terms of probability distributions.
Assume we have prior knowledge in the form of a probability distribution, $f_{\mbox{\small prior}}(\bt)$, and measured data, $\vc y$, which is believed to be a realization of the model.
The posterior knowledge, combining prior knowledge with the one acquired from data, is calculated by means of eq.
\begin{equation}\label{Bayes}
  f_{\mbox{\small post}}(\bt|\vc y)
  =
  \frac{f_{\mbox{\small prior}}(\bt)L(\vc y|\bt)}
  {\int f_{\mbox{\small prior}}(\bt')L(\vc y|\bt')d\bt'}\,,
\end{equation}
where $L(\vc y|\bt)$ is the so-called {\em likelihood function}, that is, the probability distribution, for model outputs given model parameters, evaluated at the measured data.
Here, we assume prior knowledge to be that parameters must be non-negative, but otherwise we assume the prior influence to be negligible. That is, we assume the posterior to be data-driven.

Before we set out to derive the likelihood function, from the model equations (\ref{sde}) through (\ref{data}), let us express the parameters and state variable with dimensionless quantities.
The noise parameter $\gamma$ is already dimensionless due to scale-invariance of the noise term.
We replace the state variable $S(t)$ and parameter $K$ by dimensionless quantities $q(t)$ and $\beta$, respectively, by means of the transformations
\begin{equation}
  \beta=\sqrt{\frac{T\gamma}{K}}\,,\quad
  S(t)=\frac{T\gamma r(t)}{\beta^2}e^{\beta q(t)}\,.
\end{equation}
W.r.t. these new  variables and parameters, the model equation (\ref{sde}) becomes
\begin{equation}
  \dot q(t)
  =
  \frac{\beta}{T\gamma}e^{-\beta q(t)}
  -
  \frac{1}{T}\rho(t)
  +
  \frac{1}{\sqrt{T}}\eta(t)\,,
\end{equation}
with
\begin{equation}
  \rho(t)
  =
  \frac{T}{\beta}\frac{\mbox d}{\mbox{dt}}\left[ \ln \left( r(t) \right) \right]
  +
  \frac{(2+\gamma)\beta}{2\gamma}\,.
\end{equation}
For our algorithm it is important to have the model equation in a form where the noise term does neither depend on the state variables nor the parameters that have to be inferred. In a one dimensional model this can always be achieved through re-parametrization.

The probability $P(q_1,T|q_0,0)$ of finding the system in a state $q_1$ at time $t = T$ given that it was in an initial state $q_0$ at time $t = 0$, is expressed in the form of a {\em path-integral} as
\begin{equation}\label{pathint}
P(q_1,T|q_0,0)
=
\frac{1}{Z}
\int
e^{-{\mathcal S}[q,\dot q]}
\delta(q(T)-q_1)
\delta(q(0)-q_0)
\mathcal{D}q \,,
\end{equation}
where the integral extends over all paths $q:[0,T]\rightarrow \mathbb R$.
The path-measure $\mathcal Dq$ is formally written as the infinite product
\begin{equation}\label{pathmeasure_q}
{\mathcal Dq}
=
\prod_{t}
dq(t)\,.
\end{equation}
The {\em action} is a functional on the space of paths and reads as \cite{lau_2007}
\begin{equation}\label{action1}
{\mathcal S}[{q},\dot q]
=
\frac{1}{T}
\int_0^T \dt \left\{
\frac{1}{2}
\left(
    T\dot q(t)
    +
    \rho(t)
    -
    \frac{\beta}{\gamma}e^{-\beta q(t)}\right)^2
    -
    \frac{\beta^2}{2\gamma}e^{-\beta q(t)}
\right\} \,.
\end{equation}
Note that the action includes the Jacobian that arises when changing coordinates from ${\eta(t)}$ to $q(t)$.
We introduce the time-dependent {\em Hamiltonian}
\begin{equation}\label{H}
  \mathcal{H}(q,t)
  =
  \frac{1}{\gamma}e^{-\beta q}+q\rho(t)\,,
\end{equation}
and rewrite action~(\ref{action1}) as,
\begin{eqnarray}\label{action}
{\mathcal S}[{q},\dot q] \nonumber
\\
= \frac{1}{T}
\int_0^T dt\left\{
    \frac{1}{2}
    T^2\dot q^2(t) +
    \frac{1}{2}
    \left(\rho(t)-\frac{\beta}{\gamma}e^{-\beta q(t)}\right)^2 -
    T\frac{\partial \mathcal{H}(q,t)}{\partial t} -
    \frac{\beta^2}{2\gamma}e^{-\beta q(t)}
     %\frac{(2+\gamma)\beta^2}{4\gamma}
\right\} \nonumber
\\
+ \mathcal{H}(q(T),T) - \mathcal{H}(q(0),0) \nonumber
\\
= \frac{1}{T}
\int_0^T dt\left\{
    \frac{1}{2}
    T^2\dot q^2(t) +
    \frac{1}{2}
    \left(\rho(t)-\frac{\beta}{\gamma}e^{-\beta q(t)}\right)^2 -
    Tq(t)\dot\rho(t) -
     \frac{\beta^2}{2\gamma}e^{-\beta q(t)}
%    \frac{(2+\gamma)\beta^2}{4\gamma}
\right\}  \nonumber
\\
+
    \frac{1}{\gamma}e^{-\beta q(T)}+q(T)\rho(T)
   -\frac{1}{\gamma}e^{-\beta q(0)}-q(0)\rho(0)
\,.
\end{eqnarray}

Properties of (a transformed version of) (\ref{sde}) have been derived in \cite{dutre_1977_SDE, fujisaka_1986_intermittency} and we do not repeat them here, as the focus of this paper is not on the properties of this particular model. However, in order to prove two claims made earlier, we calculate the {\em equilibrium distribution} $P_{\eq}(q) = \lim_{T\rightarrow\infty} P(q,T|q_0,0)$ in the simple case of a constant input $r(t)\equiv r_{0}$. After plugging (\ref{pathint}) and (\ref{action}) into the detailed balance condition,
\begin{equation}\label{detailed_balance}
P(q_1 t_1 | q_0 t_0 ) P_{\eq}(q_0) = P(q_0 t_1 | q_1 t_0 ) P_{\eq}(q_1) \,,
\end{equation}
and using transformation $q(t) \rightarrow q(-t)$, which maps paths from $q_0$ to $q_1$ onto paths from $q_1$ to $q_0$, we get, since $\dot\rho(t)= 0$,
\begin{equation}\label{Peq}
  P_{\eq}(q)
  \propto
  e^{-2\mathcal{H}(q)}\,.
\end{equation}
Transforming back to the original variables, it turns out that $P_{\eq}(S)$ is an inverse gamma distribution with scale parameter $2Kr_{0}/\gamma$ and shape parameter $(2+\gamma)/\gamma$, i.e.,
\begin{equation}\label{inverse_gamma}
  P_{\eq}(S)
  \propto
  S^{-2(1+\gamma)/\gamma}e^{-2Kr_{0}/(\gamma S)}\,,
\end{equation}
whose mean equals the equilibrium solution of the unperturbed system ($\gamma=0$),
\begin{equation}\label{equilibrium_mean}
  \langle S\rangle_{\eq}=Kr_{0},
\end{equation}
and whose variance, for $\gamma< 2$, is given by
\begin{equation}
  \langle (S - \langle S\rangle_{\eq})^2\rangle_{\eq}
  =
  K^2r_{0}^2
  \frac{\gamma}{2-\gamma}\,
\end{equation}
and diverges, for $\gamma\geq 2$.
The power-law decay of the inverse gamma distribution is reminiscent of the scale-invariance of the error model.

If we denote the parameter vector $\bt=(\beta,\gamma)^T$ and assume a flat prior, the posterior (\ref{Bayes}), as a function of $\bt$, is proportional to the likelihood function, which is expressed as the path-integral
\begin{equation}\label{posterior_pathint}
  f_{\mbox{\small post}}(\bt | \vc y)
  \propto
  \int
  \exp\bigg[
    -\frac{1}{2}
    \sum_{s=1}^{n+1}
    \frac
    {(\ln(y_s/r(t_s))-{\beta q(t_s)})^2}
    {\sigma^2} -{{\mathcal S}}[{q},\dot q]
    %-\ln(K\gamma)
  \bigg]
  {\mathcal Dq}
  \,.
\end{equation}
The first term in the exponent describes the log-probability distribution of model outputs, given model parameters, inputs and a system realization $q(t_s)$. The second term is the log-probability of a system realization $q(t)$ and the path-integral extends over all possible system realizations. Instead of undertaking an often prohibitive numerical computation of such integral, we apply a Hamiltonian Monte Carlo method, outlined in the next section, to sample parameter vectors from a joint distribution of system realizations and model parameters given by an appropriate discretization of the action of the path-integral.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Inference Algorithm}

In order to derive an efficient algorithm to draw parameter samples from (\ref{posterior_pathint}) we interpret it as the partition function of a 1D statistical mechanical system and simulate the dynamics of the latter, employing the so-called {\em Hamiltonian Monte Carlo} (HMC) algorithm \cite{duane_1987}.
The model parameters $\bt$ are interpreted as additional dynamical degrees of freedom coupling to the system variables $q(t)$.
Each degree of freedom, $q(t)$ and $\bt$ in our case, is paired with a conjugate variable, $p(t)$ and ${\pmb\pi}$, respectively, and the system is defined by the  Hamiltonian
\begin{equation}\label{Hamiltonian}
    \mathcal{H}_{\mbox{\tiny HMC}}(q,\bt; p,{\pmb\pi})
    =
    K( p,{\pmb\pi}) + V( q,\bt)\,,
\end{equation}
where
\begin{equation}\label{K}
   K( p,{\pmb\pi})
   =
   \int_0^T \frac{ p^2(t)}{2m(t)}dt
   + \sum_{\alpha=1}^2\frac{\pi_\alpha^2}{2m_\alpha}\,,
\end{equation}
and $V( q,\bt)$ is the negative logarithm of the kernel of (\ref{posterior_pathint}).
The posterior (\ref{posterior_pathint}) can then be expressed by the phase space path integral
\begin{equation}\label{phaseSpacePathInt}
    f_{\mbox{\small post}}(\bt | \vc y)
  \propto
  \int
  e^{-\mathcal {H}_{\mbox{\tiny HMC}}(q,\bt; p,{\pmb\pi})}
  {\mathcal Dp}
   {\mathcal Dq}
   d{\pmb\pi}
  \,.
\end{equation}

The HMC method, which is a combination of the {\em Metropolis algorithm} \cite{metropolis_1953} and {\em molecular dynamics} methods \cite{alder_1959_MD, rahman_1964_MD}, iterates the following steps:
\begin{enumerate}
  \item
  Momenta $p(t)$ and $\pi$ are sampled from the Gaussian distributions defined by Eq.~(\ref{K}).
  \item
  The system is then allowed to evolve in $\left(q,\bt; p,{\pmb\pi}\right)$-phase space for an arbitrary time interval $\tau$ according to a volume-preserving and time-reversible solution of a discretization of Hamilton's equations.
  %The system is integrated by means of a volume-preserving and time-reversible %solution of a discretization of Hamilton's equations.
  \item
  The discretization error on the energy preservation due to the previous step is corrected by a Metropolis acceptance/rejection step.
\end{enumerate}
The last step is the standard Metropolis algorithm, while the first two steps allow us to make arbitrarily large jumps in phase space while maintaining an arbitrarily large acceptance rate. Each new phase space configuration is associated with a combination of model parameters $\bt$, which is compatible with the data in the Bayesian sense. Thus, the parameter marginal of the simulated Markov chain of configurations represents a sample of the posterior probability distribution.

In order to simulate the dynamics of Hamiltonian (\ref{Hamiltonian}), we first need to discretize path-integral (\ref{phaseSpacePathInt}).
Therefore, let us assume that the measurement time points $\left\{ y_s \right\}_{s=1,\dots, n+1}$ of the time series (\ref{data}) are equidistantly distributed on the time interval $[0,T]$, with $t_1=0$ and $t_{n+1}=T$.
Each interval between two consecutive data points is further partitioned into $j$ bins, such that we have a total of $nj+1=N>>1$ discretization points.
The path-integral (\ref{phaseSpacePathInt}) is then approximated by an ordinary integral, with the approximate path-measure
\begin{equation}
  \mathcal Dp\mathcal Dq
  \approx
  \prod_i dp_i dq_i
\end{equation}
and the discretized versions of $K( p,{\pmb\pi})$ and $V( q,\bt)$ are given by eqs.
\begin{eqnarray}
   K( p,{\pmb\pi}) \approx
   \sum_{i=1}^N
   \frac{ p_i^2}{2m_i}\Delta t
   +
   \sum_{\alpha=1}^2\frac{\pi_\alpha^2}{2m_\alpha}\,,\label{Kdisc}
   \\
   V(q,\bt) \approx \frac{\Delta t}{T} \sum_{i=2}^{N}
   \left\{ \frac{1}{2} T^2 \dot q_i^2 + \frac{1}{2}
     \left( \rho_i-\frac{\beta}{\gamma}e^{-\beta q_i} \right)^2 -
    \frac{\beta^2}{2\gamma} e^{-\beta q_i} - T q_i\dot\rho_i \right\}  \nonumber
  \\
  +
  \frac{1}{\gamma}
  e^{-\beta q_N}
  +
  q_N \rho_{N}
  -
  \frac{1}{\gamma}
  e^{-\beta q_1}
  -
  q_1 \rho_{1}
  \nonumber
  \\
  +
  \sum_{s=1}^{n+1}
  \frac{(\ln(y_s/r_{(s-1)j+1}) - {\beta q_{(s-1)j+1}})^2}{2\sigma^2}
  \label{Vdisc}
   \,,
\end{eqnarray}
with
\begin{equation}
  \dot q_i = \frac{q_i-q_{i-1}}{\Delta t}\,,
  %\quad
%  \bar q_i = \frac{q_i+q_{i-1}}{2}\,,
\end{equation}
and
\begin{equation}\label{rhodisc}
\rho_i = \frac{T}{\beta} \frac{\ln(r(t_{i})/r(t_{i-1}))}{\Delta t}
+
\frac{(2+\gamma)\beta}{2\gamma}
\,,\quad
\dot\rho_i = \frac{\rho_i-\rho_{i-1}}{\Delta t}\,.
\end{equation}
Note that we've neglected terms of order $\mathcal O(N^{-1/2})$ in the action (\ref{Vdisc}).

The discretized Hamiltonian describes a {\em classical polymer chain} of $N$ beads with harmonic bonds between neighbouring beads in an external field. The latter consists of two parts, a field that results from the dynamics of the original Eq. (\ref{sde}) and is felt by all the beads, and a field that results from the measurements and is felt by the measurement beads only. The masses $m_i$ and $m_{\alpha}$ are tunable parameters of the algorithm.
Since measurement beads are constrained more than intermediate beads we will assign larger masses to the former.
Fig. (\ref{fig:polymer}) shows a typical realization of the dynamics of the polymer. 
The physical time (horizontal axis) is interpreted as a spatial dimension while a fictitious simulation time is introduced (vertical axis).
Measurements beads only move within the measurement uncertainty whilst the intermediate beads explore much larger regions of phase space.

%The dynamics of the system will be therefore mostly dependent on the lighter intermediate particles, while the heavy measurement beads will be constrained in the vicinity of the measured data points. We thus reduce the original Bayesian inference problem to simulating the dynamics of a linear polymer, as sketched in Fig.~(\ref{fig:polymer}). Each \emph{state} of this fictitious molecule corresponds to a well-defined configuration in the original phase space, characterized by a set of system variables $\left\{q_i\right\}_{i=1,\dots,N}$ and a parameter vector $\bt$.

\begin{figure}[htb!]
    \centering
    \captionsetup{justification=justified}
    \includegraphics[width=0.7\textwidth]{Figs/FigPolymer.png}
    \caption{Simulated dynamics of a polymer chain with $n+1=11$ data points (large filled circles) and $j-1=9$ intermediate beads (small empty circles). All other parameters are discussed in Section~\ref{numerical_results}. Top: initial state. The intermediate beads represent a linear interpolation to the data points. Bottom: polymer configuration after 1000 iterations of the propagation algorithm. The dotted line represents the initial configuration. It is evident that the new configuration is mostly determined by the dynamics of the low-mass intermediate beads, while the heavy-mass data points barely move from their initial positions.}
    \label{fig:polymer}
\end{figure}

The potential (\ref{Vdisc}) contains terms with different scaling in the potentially large numbers $N$ and $n$, which describe dynamics on different time-scales.
In particular, for large $N$, $V(q,\bt)$ is dominated by its harmonic part and a brute force numerical integration of Hamilton's equations in step ($ii$) of the HMC algorithm would require a very small discretization time-step to sufficiently resolve its dynamics.
An interesting approximative approach would be to employ a partial averaging of the fast Fourier modes as described in \cite{doll_1985_fourier}.
We choose an exact approach and employ a multiple time scale integration based on Trotter's formula \cite{tuckerman_1993}.

For this purpose it proves to be useful to introduce so-called \emph{staging} variables \cite{tuckerman_1992}, and diagonalize the harmonic part in between the measurement points.
Therefore, we rewrite the discretized harmonic part of the action as
\begin{eqnarray}
  \sum_{i=2}^{N}
  \frac{T}{2\Delta t}
  (q_i-q_{i-1})^2 \nonumber
  \\ =
  \frac{T}{2}
  \sum_{s=1}^{n}\left\{
    \frac{(q_{(s-1)j+1} - q_{sj+1})^2}{j\Delta t}
    +
    \sum_{k=2}^j
    \frac{k}{(k-1)\Delta t}
    (q_{(s-1)j+k}-q^*_{(s-1)j+k})^2
  \right\}\,,\nonumber
  \\
\end{eqnarray}
with
\begin{equation}
  q^*_{(s-1)j+k}
  =
  \frac{(k-1)q_{(s-1)j+k+1} + q_{(s-1)j+1} }{k}
  \,.
\end{equation}
We apply the coordinate transformations
\begin{equation}
  u_{sj+1} = q_{sj+1}\,, \qquad\qquad s=0,\dots,n\, ,
\end{equation}
for the boundary beads, corresponding to the original measurement points, and
\begin{equation}
  u_{sj+k} = q_{sj+k} - q^*_{sj+k}\,, \quad s=0,\dots,n-1\, , \quad k=2,\dots,j\, ,
\end{equation}
for the intermediate staging beads.
The inverse transformations are given by
\begin{eqnarray}
  q_{sj+1} = u_{sj+1}\,,
  \\
  q_{sj+k} = \sum_{l=k}^{j+1}\frac{k-1}{l-1}u_{sj+l}
  +\frac{j-k+1}{j}u_{sj+1}\,.
\end{eqnarray}
The latter can be equivalently expressed by the recursive relation
\begin{equation}
  q_{sj+k} = u_{sj+k} + \frac{k-1}{k} q_{sj+k+1}+ \frac{1}{k}u_{sj+1} \,.
\end{equation}

We split the Hamiltonian $\mathcal{H}_{\mbox{\tiny HMC}}$ into components with different scaling behaviour in the potentially large numbers $n$ and $N$ and write
\begin{equation}
  \mathcal{H}_{\mbox{\tiny HMC}}= \mathcal H_N+\mathcal H_n+\mathcal H_1\,,
\end{equation}
where
\begin{eqnarray}
  \mathcal H_N =
  \frac{1}{2}
  \sum_{s=1}^{n}
  \sum_{k=2}^j
  \left\{
    \frac{\Delta t}{m'}p_{(s-1)j+k}^2
    +
    \frac{Tk}{\Delta t(k-1)}
    u_{(s-1)j+k}^2
  \right\}\,,\label{H_N}
  \\
  \mathcal H_n =
  \frac{1}{2}
  \sum_{s=1}^{n+1}
  \left\{
   \frac{\Delta t }{M}p_{(s-1)j+1}^2
    +
    \frac{(\ln(y_s/r_{(s-1)j+1}) - {\beta u_{(s-1)j+1}})^2}{\sigma^2}
   \right\} \nonumber
   \\
  +
  \frac{T}{2j\Delta t}
  \sum_{s=1}^{n}
    (u_{(s-1)j+1} - u_{sj+1})^2
   \,, \label{H_n} \\
  \mathcal H_1=
   \sum_{\alpha=1}^2\frac{\pi_\alpha^2}{2m_\alpha}
   +
  \frac{\Delta t}{T}
   \sum_{i=2}^{N}
   \left\{
    \frac{1}{2}
     \left(
        \rho_i-\frac{\beta}{\gamma}e^{-\beta q_i}
     \right)^2
    -
    \frac{\beta^2}{2\gamma}
    e^{-\beta q_i}
   -
    T q_i\dot\rho_i
   \right\} \nonumber
  \\
  +
  \frac{1}{\gamma}
  e^{-\beta q_N}
  +
  q_N \rho_{N}
  -
  \frac{1}{\gamma}
  e^{-\beta q_1}
  -
  q_1 \rho_{1} \,. \label{H_1}
\end{eqnarray}
We have introduced two different masses, $M$ and $m'$, for the boundary and staging beads, respectively. 
The harmonic part for the staging beads given in Eq.~(\ref{H_N}) scales linearly with $N$. The terms of Eq.~(\ref{H_n}), including both the harmonic part for the boundary beads and the measurement term, scale linearly with $n$. Finally, Eq.~(\ref{H_1}) does not scale with neither $n$ nor $N$.
Thanks to the staging variables, $\mathcal H_N$ decouples completely from $\mathcal H_n$.

We use the Trotter formula according to \cite{tuckerman_1992} in order to design a reversible molecular dynamics integrator that takes these different time scales into account.
In order to design an appropriate partition of the Hamiltonian, we need to distinguish different regimes, such as
\begin{enumerate}
  \item[\it i.]
  $\mathcal H_N \sim \mathcal H_n >> \mathcal H_1$,
  \item[\it ii.]
  $\mathcal H_N >> \mathcal H_n \sim \mathcal H_1$,
  \item[\it iii.]
  $\mathcal H_N >> \mathcal H_n >> \mathcal H_1$.
\end{enumerate}
Here, we restrict ourselves to regime ($ii$), i.e., we assume that the number of measurements $n$ is not too large and/or the measurement error $\sigma$ is not too small.
The generalization of the method to the other schemes is straightforward.
In regime ($ii$) we simply separate the harmonic part of the action, for the staging beads, from the rest and write
\begin{equation}
  \mathcal{H}_{\mbox{\tiny HMC}}=\mathcal H_N + \mathcal H'\,.
\end{equation}
In order to design a reversible integrator for the associated Hamilton equations, we define the Liouville operators
\begin{equation}
  iL_N=\{\cdot\,,\,\mathcal H_N\}\,,\quad
  iL'=\{\cdot\,,\,\mathcal H'\}\,,
\end{equation}
where $\{\cdot\,,\,\cdot\}$ denotes the Poisson brackets that are defined on functions on the phase space.
Trotter's formula \cite{trotter_1959} allows us to write the Hamiltonian propagator as
\begin{equation}\label{propagator}
  e^{i(L_N+L')\tau}
  =
  (e^{iL_N(\Delta\tau/2)}e^{iL'\Delta\tau}e^{iL_N(\Delta\tau/2)})^P
  +
  \mathcal O(\tau^3/P^{2})\,,
\end{equation}
for $\tau =P\Delta \tau$.
In regime ($ii$) the outer propagator $\exp[iL_N(\Delta \tau/2)]$ describes much faster dynamics than the inner one. However, it is the dynamics of uncoupled harmonic oscillators, which we can readily solve.
Masses and frequencies of the oscillators are derived from (\ref{H_N}) as
\begin{equation}
  m=m'/\Delta t\,,\quad
  \omega_k=\sqrt{\frac{Nk}{(k-1)m}}\,.
\end{equation}
The fast outer propagator is then given by the equations,
\begin{eqnarray}
  u_{(s-1)j+k}(\Delta\tau/2) \nonumber \\
  = u_{(s-1)j+k}(0)\cos(\omega_k\Delta\tau/2)
  +
  \frac{p_{(s-1)j+k}(0)}{m\omega_k}\sin(\omega_k\Delta\tau/2)\,,\\
  p_{(s-1)j+k}(\Delta\tau/2) \nonumber \\
  = p_{(s-1)j+k}(0)\cos(\omega_k\Delta\tau/2)
  -
  m\omega_k u_{(s-1)j+k}(0) \sin(\omega_k\Delta\tau/2)\,,
\end{eqnarray}
for $s=1,\dots,n$ and $k=2,\dots,j$.

For the inner, slow propagator, we employ the velocity Verlet algorithm \cite{swope_1982_verlet}.
For the boundary beads, it reads
\begin{eqnarray}
  u_{(s-1)j+1}(\Delta\tau) \nonumber \\
  = u_{(s-1)j+1}(0)
  +
  \frac{\Delta\tau}{M} p_{(s-1)j+1}(0)
  +
  \frac{\Delta \tau^2}{2M}
  F_{(s-1)j+1}[\vc u(0),{\pmb\theta}(0)]\,,\\
  p_{(s-1)j+1}(\Delta\tau) \nonumber \\
  = p_{(s-1)j+1}(0)
  +
  \frac{\Delta\tau}{2}
  (
  F_{(s-1)j+1}[\vc u(0),{\pmb\theta}(0)]
  +
  F_{(s-1)j+1}[\vc u(\Delta\tau),{\pmb\theta}(\Delta\tau)]
  )\,,
\end{eqnarray}
with $s=1,\dots,n+1$ and where $F_i[\vc u,{\pmb\theta}]$ denotes the partial derivative of $\mathcal H'[\vc u,{\pmb\theta}]$ w.r.t. $u_i$.
For the model parameters $\bt$ and their momenta ${\pmb\pi}$, analogous equations apply.
When it comes to the staging beads, however, only the momenta have to be updated, because the associated kinetic term is not part of $\mathcal H'$ but of $\mathcal H_N$.
Thus,
\begin{eqnarray}
  p_{(s-1)j+k}(\Delta\tau) \nonumber \\
  =
  p_{(s-1)j+k}(0)
  +
  \frac{\Delta\tau}{2}
  (
  F_{(s-1)j+k}[\vc u(0),{\pmb\theta}(0)]
  +
  F_{(s-1)j+k}[\vc u(\Delta\tau),{\pmb\theta}(\Delta\tau)]
  )
  \,,
\end{eqnarray}
with $s=1,\dots,n$ and $k=2,\dots,j$. The above propagators are applied sequentially $P$ times to calculate the system evolution over the time $\tau$.
After $P$ steps, the proposed configuration, $(\vc u',\bt';\vc p',{\pmb\pi}')$ is accepted with the Metropolis probability
\begin{equation}
  \min
  \left(
    1,
    e^{\mathcal{H}_{\mbox{\tiny HMC}}(\vc u,\bt;\vc p,{\pmb\pi}) - \mathcal{H}_{\mbox{\tiny HMC}}(\vc u',\bt';\vc p',{\pmb\pi}')}
  \right)\,,
\end{equation}
and the next iteration starts with sampling an new momentum vector $(\vc p,{\pmb\pi})$.

\section{Numerical Results}\label{numerical_results}

The algorithm described above was implemented using the free and open-source programming language \emph{Julia} (version 0.3.9)~\cite{julia_web}. One of the attractive features of the language is its (fast growing) library of over 600 packages for numerical and scientific applications.
In our case, we made extensive use of the \emph{ReverseDiffSource} package, which provides a powerful tool for reverse-mode automated differentiation (AD) of arbitrarily complex user-defined expressions. Our algorithm benefits greatly from the use of AD. Indeed, it gives us the possibility to modify Eq.~(\ref{sde}) and therefore the action~(\ref{action}) leaving the implementation of the algorithm unaltered. This clearly gives our program significant flexibility making it suitable for a much broader range of applications than the simple exemplary SDE model described here.
%
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\textwidth]{Figs/FigRainData.png}
    \caption{\label{fig:rain_data_S}Sinusoidal input $r(t)$ (dashed line), "true" system realization with parameters $K_{\mbox{\small true}} = 50$ (in some arbitrary units of time) and $\gamma_{\mbox{\small true}} = 0.2$ (solid line) and synthetic observed data. One may notice how the system response follows the oscillations of the input signal.}
\end{figure}

We have run preliminary simulations on a 64-bit laptop computer equipped with a Windows 7 operating system, a double-core 1.6~GHz CPU (Intel i5-4200U) and 8~GB of RAM. We have used a serial implementation of the algorithm. We assumed $n+1 = 11$ measurement points and $j = 10, 20,\dots, 50$ intermediate staging beads, with a total number of discretization points $N = 101, 201, \dots, 501$. We have set $\Delta\tau = 0.25$ and $P = 3$ in the Hamiltonian propagator~(\ref{propagator}), with a constant total observation time $T = 833$ (arbitrary units of time). The "true" values of the parameters to be inferred were set to $K_{\mbox{\small true}}=50$ (arbitrary units of time) and $\gamma_{\mbox{\small true}} = 0.2$. Their initial values in the simulations were set to $K=200$ and $\gamma = 0.5$.
%
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\textwidth]{Figs/FigSpaghetti.png}
    \caption{Simulated system realizations and corresponding synthetic data generated as described in the text with $j=30$ staging beads and therefore $N=301$ discretization points. In the inset one may appreciate the different dynamics of heavy data points and light staging beads.}
    \label{fig:spaghetti}
\end{figure}

We have assumed a simple sinusoidal input $r(t) = \sin^2 \left( 0.01 t \right) + 0.1$. A system realization was first obtained from Eq.~(\ref{sde}) using $K_{\mbox{\small true}}$ and $\gamma_{\mbox{\small true}}$. Such system realization was then used to generate a synthetic time series of observed data according to Eq.~(\ref{data}). The error $\sigma$ was set to $0.1$. The input signal, the "true" system realization and the corresponding data time series are shown in Fig.~(\ref{fig:rain_data_S})

Within the investigated range of parameters we observed a reasonably linear dependence of the total computing time on the number of discretization points $N$. For example, with $j=30$ (i.e., $N = 301$ discretization points) a complete run with 20000 iterations required about 13 minutes. A set of 200 simulated system realizations with $j=30$ is shown in Fig.~(\ref{fig:spaghetti}) together with the synthetic data generated as described above. In the simulation the masses were set (in arbitrary units) to $M=720$, for the boundary beads, $m=130$, for the staging beads and $m_\alpha=150$, for both the dimensionless parameters $\beta$ and $\gamma$. The different dynamics of the heavy boundary beads and the light staging beads can be appreciated in Fig.~(\ref{fig:spaghetti}).
%
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\textwidth]{Figs/FigChainK.png}
    \caption{Markov chain for the inferred parameter $K$.}
    \label{fig:chainK}
\end{figure}
%
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\textwidth]{Figs/FigChainG.png}
    \caption{Markov chain for the inferred parameter $\gamma$.}
    \label{fig:chainG}
\end{figure}

The Markov chains for the parameters $K$ and $\gamma$ generated after 50000 iterations of the HMC algorithm are shown in Figs.~(\ref{fig:chainK}) and (\ref{fig:chainG}), respectively. Both of them are clearly compatible with the true parameter values $K_{\mbox{\small true}} = 50$ and $\gamma_{\mbox{\small true}}=0.2$.
The efficiency of the algorithm can be best appreciated by inspecting the system evolution in the phase space $K-\gamma$, as shown in Fig.~(\ref{fig:phase_space_evol}). The very first step of the algorithm already takes the system to the vicinity of the true parameter values, where most of its dynamics then occurs. Few excursions far away from the true parameter values explore the heavy tails of the posterior parameter distribution.
%\begin{figure}[htb!]
%    \centering
%    \includegraphics[width=0.3\textwidth]{Figs/FigFinalState.png}
%    \caption{PDF for the system final state expressed via the coordinate set $\left\{ q_N \right\}$. The dashed curve represents a normal distribution with the same mean and variance. The final state is basically normally distributed.}
%    \label{fig:final_state}
%\end{figure}
%
\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\textwidth]{Figs/FigPhaseSpaceEvol.png}
    \caption{System dynamics in the phase space $K-\gamma$. The empty circle represents the initial state, while the square corresponds to the true parameter values used to generate the data.}
    \label{fig:phase_space_evol}
\end{figure}

%\begin{figure}[htb!]
%    \centering
%    \includegraphics[width=0.7\textwidth]{Figs/FigK.png}
%    \caption{Probability density function for the inferred parameter $K$. The true value used to generate the data, $K_{\mbox{\small true}}=50$, is represented by the dotted vertical line.}
%    \label{fig:K_distr}
%\end{figure}
%
%\begin{figure}[htb!]
%    \centering
%    \includegraphics[width=0.65\textwidth]{Figs/FigG.png}
%    \caption{Probability density function for the inferred parameter $\gamma$. The true value used to generate the data, $K_{\mbox{\small true}}=50$, is represented by the dotted vertical line.}
%    \label{fig:G_distr}
%\end{figure}
The Markov chains of Figs.~(\ref{fig:chainK}) and (\ref{fig:chainG}) naturally lead to the probability density functions (PDF) for the two parameters of interest, $K$ and $\gamma$, respectively. These functions were calculated using the built-in kernel density estimator provided by Mathematica (version 10) and are shown in Fig.~(\ref{fig:KG_distr}).
\begin{figure}[htb!]
    \centering
    \includegraphics[width=1.0\textwidth]{Figs/FigKg.png}
    \caption{Probability density functions for the inferred parameters $K$ (left) and $\gamma$ (right). The true values used to generate the data, $K_{\mbox{\small true}}=50$ and $\gamma_{\mbox{\small true}}=0.2$, are represented by the dotted vertical lines.}
    \label{fig:KG_distr}
\end{figure}

The structure of the HMC algorithm is well suited to be parallelized. 
For hydrological applications, where the number of measurement points are often in the tens of thousands, this is an important advantage.

%We tested the built-in Julia multiprocessing environment to run the program on two separate processes, but we observed a much slower performance compared to the serial implementation. This should not be surprising since with the small number of discretization points used so far, not more than a few hundreds, the overheads associated with the communication between the parallel processes significantly exceeded their advantages. Nevertheless, the method might be applied to systems requiring a much larger number of points. In hydrology, for instance, typical time series corresponding to hourly observations of precipitations and river discharges may contain several tens of thousands of measurements. In such cases a parallel version of the program would definitely improve the performance of the method. However, it is interesting to observe that Julia might not be the best option for parallelization. Indeed, the multiprocessing environment provided by Julia is strongly task oriented, while our problem is fundamentally data oriented. In view of a future parallelization of the HMC algorithm it is therefore advisable to investigate also alternative programming languages, such as C++. This issue will be discussed in a forthcoming publication.

\section{Conclusions}

We've presented a new algorithm, for the generation of posterior parameter samples, for ordinary 1D SDE models that are calibrated to time-series.
The algorithm is derived from a re-interpretation of the posterior distribution as a partition function of a 1D statistical mechanical system and employs a Hamiltonian Monte Carlo approach combined with a multiple time-scale integration.
Furthermore, a generic re-parametrization is suggested, which, to some extent, decouples harmonic modes in between measurement points from both the measurement points and the model parameters, and allows for an efficient analytical solution of their dynamics.
The algorithm can be implemented in an efficient and generic fashion using automated differentiation and parallelization.
Furthermore, it can easily be adapted to other inference problems.
In particular, it can be adapted to higher dimensional SDEs and SDEs coupled to ODEs.
We explore these adaptations in our future work.

\section*{References}
\bibliographystyle{unsrt}
\bibliography{refs}

\end{document}